Enzo+Moray: Radiation Hydrodynamics Adaptive Mesh Refinement Simulations with Adaptive Ray Tracing

### 2011-02-09 mnras/mnras.tex
Massive figure filename renaming because of figure number changes with the new figures.
Proofread.

\begin{figure}
  \plotone{test1_pdf.eps}
  \caption{\label{fig:test1_pdf} Test 1 (\hii region expansion with a
    monochromatic spectrum of 13.6 eV). Probability distribution
    function for neutral fraction at 10 Myr (solid), 100 Myr (dashed),
    and 500 Myr (dotted).  Recombination of hydrogren at T = $10^4$ K
    causes the maximum neutral fraction to decrease from 1 to 0.75.}
\end{figure}

\begin{figure}
  \plotone{f6.eps}
  \caption{\label{fig:test1_HI} Test 1 (\hii region expansion
    with a monochromatic spectrum of 13.6 eV). Slice of neutral
    fraction at the origin at 500 Myr.}
\end{figure}

@@

\begin{figure}
  \plotone{test2_pdf}
  \caption{\label{fig:test2_pdf} Test 2. (\hii region expansion with a
    $T=10^5$ K blackbody spectrum).  Probability distribution function
    for ionized fraction (left) and temperature (right) at 10 Myr
    (solid), 100 Myr (dashed), and 500 Myr (dotted).}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{f9}
  \epsscale{1}
  \caption{\label{fig:test2_2} Test 2. (\hii region expansion
    with a $T=10^5$ K blackbody spectrum).  Top: Slices through the
    origin of neutral fraction at 10 and 100 Myr.  Bottom: Slices of
    temperature at 10 and 100 Myr.}
\end{figure*}

\begin{figure}
  \plotone{f10}
  \caption{\label{fig:test2_3} Test 2. (\hii region expansion
    with a $T=10^5$ K blackbody spectrum).  Top: Evolution of the
    radius of the simulated ionisation front (dashed) and analytical
    (solid) Str\"{o}mgren radius.  Bottom: The ratio of the calculated
    and analytical Str\"{o}mgren radius.}
\end{figure}

@@

\begin{figure}
  \plotone{f11}
  \caption{\label{fig:test3_1} Test 3. (I-front trapping in a dense
    clump and shadowing).  Line cut from the point source through the
    middle of the dense clump at $t = 1, 3, 5, 15$ Myr.  of the
    average neutral fraction (top) and temperature (bottom) of the
    clump.}
\end{figure}

\begin{figure}
  \plotone{f12}
  \caption{\label{fig:test3_2} Test 3. (I-front trapping in a dense
    clump and shadowing).  Evolution of the average ionised fraction
    (top) and temperature (bottom) of the overdense clump.}
\end{figure}

\begin{figure}
  \plotone{test3_pdf}
  \caption{\label{fig:test3_pdf} Test 3. (I-front trapping in a dense
    clump and shadowing). Probability distribution function for
    ionized fraction (left) and temperature (right) at 1 Myr (solid),
    3 Myr (dashed), and 15 Myr (dotted) within the dense clump.}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{f13}
  \epsscale{1}
  \caption{\label{fig:test3_3} Test 3. (I-front trapping in a dense
    clump and shadowing).  Clockwise from upper left: Slices through
    the origin of neutral fraction (1 Myr), temperature (1 Myr),
    temperature (15 Myr), and neutral fraction (15 Myr).}
\end{figure*}

@@

\begin{figure}
  \plotone{test4_pdf}
  \caption{\label{fig:test4_pdf} Test 4. (Multiple cosmological
    sources). Probability distribution function for neutral fraction
    (left) and temperature (right) at 50 kyr (solid) and 200 kyr
    (dashed).}
\end{figure}

\begin{figure}
  \plotone{f14}
  \caption{\label{fig:test4_1} Test 4. (Multiple cosmological
    sources).  Evolution of the mass- (dashed) and volume-averaged
    (solid) ionised fraction.}
\end{figure}

\begin{figure*}
  %\epsscale{0.9}
  \epsscale{2}
  \plotone{f15}
  \epsscale{1}
  \caption{\label{fig:test4_2} Test 4. (Multiple cosmological
    sources).  Top: Slices through the origin of neutral fraction at
    50 and 200 kyr at the coordinate $z = z_{\rm box}/2$.  Bottom:
    Slices of temperature at 50 and 200 kyr.  No smoothing has been
    applied to the images.}
  %\epsscale{1}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.8}
  \plotone{f16}
  \caption{\label{fig:test5_1} Test 5. (\hii region in a uniform
    medium).  Top: Growth of the computed ionisation front radius at
    an ionised fraction $x_e = 0.5$ (dashed) and at a temperature $T =
    10^4$ K (dotted) compared to the analytical estimate [solid;
    equation (\ref{eqn:test5_r})].  Bottom: The ratio of the computed
    ionisation front radii to the analytical estimate.} 
  %\epsscale{1}
\end{figure}

\begin{figure}
  \plotone{f17}
  \caption{\label{fig:test5_2} Test 5. (\hii region in a uniform
    medium).  Clockwise from the upper left: Radial profiles of
    density, temperature, ionised fraction, and pressure at times $t
    =$ 10, 200, and 500 Myr.}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{f18}
  \epsscale{1}
  \caption{\label{fig:test5_3} Test 5. (\hii region in a uniform
    medium).  Clockwise from the upper left: Slices through the origin
    of ionised fraction, neutral fraction, temperature, and density at
    time $t =$ 500 Myr.}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.8}
  \plotone{f19}
  \caption{\label{fig:test6_1} Test 6. (\hii region in a 1/$r^2$
    density profile).  Top: Growth of the computed ionisation front
    radius as $T=10^4$ K and $x_e = 0.5$ as definitions for the front.
    Bottom: Velocity of the ionisation front, computed from outputs at
    0.5 Myr intervals.  The velocity is calculated from $r_{\rm IF}$,
    whose coarse time resolution causes the noise seen in $v_{\rm
      IF}$.  It is smooth within the calculation itself.}
  %\epsscale{1}
\end{figure}

\begin{figure}
  \plotone{f20}
  \caption{\label{fig:test6_2} Test 6. (\hii region in a 1/$r^2$
    density profile).  Clockwise from the upper left: Radial profiles of
    density, temperature, ionised fraction, and pressure at times $t
    =$ 3, 10, and 25 Myr.}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{f21}
  \epsscale{1}
  \caption{\label{fig:test6_3} Test 6. (\hii region in a 1/$r^2$
    density profile).  Clockwise from the upper left: Slices through the origin
    of ionised fraction, neutral fraction, temperature, and density at
    time $t =$ 25 Myr.}
\end{figure*}

@@

For the first Myr, the radiation source creates a weak R-type front
where the medium is heated and ionised but does not expand because
$v_{\rm IF} > c_s$.  When $v_{\rm IF}$ becomes subsonic, the medium
can react to the passing ionisation front and creates a shock, leaving
behind a heated an rarefied medium.  This behavior is clearly seen in
the radial profiles of density, temperature, ionised fraction, and
pressure in Fig. \ref{fig:test6_2}.  The inner density decreases
over two order of magnitude after 25 Myr.  To illustrate any
deviations in spherical symmetry, we show in Fig. \ref{fig:test6_3}
slices of density, temperature, neutral fraction, and ionised fraction
at the final time.  The only artifact apparent to us is the slight
broadening of the shock near the $x=0$ and $y=0$ planes.  This causes
the ionisation front radius to be slightly smaller in those
directions.  In the diagonal direction, the neutral column density
through the shock is slightly smaller, allowing the high-energy
photons to photo-ionise and photo-heat the gas to $x_e = 5 \times
10^{-3}$ and $T = 2000$ K out to $\sim50$ pc from the shock.  The
reflecting boundaries are responsible for this artifact because this
is not seen when the problem is centred in the domain, removing any
boundary effects.

@@

\begin{figure}
  %\epsscale{0.9}
  \plotone{f22}
  \caption{\label{fig:test7_1} Test 7. (Photo-evaporation of a dense
    clump).  Line cuts from the point source through the middle of the
    dense clump at $t = 1, 10, 50$ Myr of (clockwise from the upper
    left) density, temperature, pressure, and neutral fraction.}
  %\epsscale{1}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{f23}
  \epsscale{1}
  \caption{\label{fig:test7_2} Test 7. (Photo-evaporation of a dense
    clump).  Clockwise from the upper left: Slices through the clump
    centre of neutral fraction, pressure, temperature, and density at
    time $t =$ 10 Myr.}
\end{figure*}

\begin{figure*}
  \epsscale{2}
  \plotone{f24}
  \epsscale{1}
  \caption{\label{fig:test7_3} Test 7. (Photo-evaporation of a dense
    clump).  Same as Fig. \ref{fig:test7_1} but at $t = 50$ Myr.}
\end{figure*}

\begin{figure}
  \plotone{test7_pdf}
  \caption{\label{fig:test7_pdf} Test 7. (Photo-evaporation of a dense
    clump).  Probability distribution function for temperature (left)
    and flow Mach number (right) at 10 Myr (solid) and 50 Myr
    (dashed).}
\end{figure}

\begin{figure}
  \plotone{test7_Mach}
  \caption{\label{fig:test7_4} Test 7. (Photo-evaporation of a dense
    clump).  Slices through the clump centre of the flow Mach number
    at t = 10 Myr (left) and 50 Myr (right).}
\end{figure}

@@

\begin{figure*}
  \epsscale{2}
  \plotone{f25}
  \epsscale{1}
  \caption{\label{fig:test8_1} Application 1. (Champagne flow from a dense
    clump).  Slices of density through the initial clump centre in the
    x-y plane (top) and x-z plane (bottom) at $t = 10, 40, 100, 150$
    kyr.  Notice the instabilities that grow from perturbations
    created while the \hii region is contained in the dense
    clump.}
\end{figure*}

@@

The $x$ and $y$-boundaries are periodic, and the $z$-boundaries are
reflecting.  These will cause artifical features, in particular,
because of the top reflecting boundary; nevertheless, these tests
provide a stress test on a radiation hydrodynamics solver.  We show
the evolution of the density, temperature, and ionised fraction of the
optically thick and optically thin cases in Figs.
\ref{fig:test9_thick} and \ref{fig:test9_thin}.  The initial state of
the Rayleigh-Taylor instability is shown in the left panels.

@@

\begin{figure}
  %\epsscale{0.75}
  \plotone{f26a}
  \plotone{f26b}
  \plotone{f26c}
  \caption{\label{fig:test9_thick} Application 2. (Irradiated Rayleigh-Taylor
    instability; optically thick case).  Slices at $y=0$ of density
    (top), temperature (middle), and electron fraction (bottom).  The
    source turns on at $t=0$.}
  %\epsscale{1}
\end{figure}

\begin{figure}
  %\epsscale{0.75}
  \plotone{f27a}
  \plotone{f27b}
  \plotone{f27c}
  \caption{\label{fig:test9_thin} Same as Fig. \ref{fig:test9_thick}
    but for the optically thin case.}
  %\epsscale{1}
\end{figure}

@@

\begin{figure*}
  \epsscale{2}
  \plotone{f28}
  \epsscale{1}
  \caption{\label{fig:test10} Application 3 (Photo-evaporation of a
    blastwave).  Slices of density (top) and temperature (bottom) at
    $t = 2.5, 3, 5, 7.5$ Myr in the $x-z$ plane.  As the R-type
    ionisation front propagates through the blastwave centre,
    instabilities grow from the slightly inhomogeneous hot and rarefied
    medium.  Note that the dense shell of the blastwave also creates
    dense inward fingers in the ionisation front shock.}
  %\epsscale{1}
\end{figure*}

@@

\begin{figure*}
  %\epsscale{1}
  \epsscale{2}
  \plotone{f29}
  \epsscale{1}
  \caption{\label{fig:test11} Application 4 (Collimated radiation from a
    dense clump).  Slices of density (top) and temperature (bottom) at
    $t = 0.1, 3.25, 10.75, 23.25$ Myr.  The conical \hii region
    drives shocks transversely into the overdense sphere and creates
    polar champagne flows.  The ambient medium is heated to $T \sim 3
    \times 10^4$ K as the ionisation front passes the
    constant-pressure cloud surface.  The ionisation front changes
    from D-type to R-type after it enters the ambient medium.}
  %\epsscale{1}
\end{figure*}

@@

\begin{figure}
  %\epsscale{1.15}
  \plottwo{f30a}{f30b}
  \caption{\label{fig:test12_kph} Application 4 (Time variations of the
    source luminosity).  Top: Slice of the photo-ionisation rate
    \kph~through the origin.  The source has a duty cycle of 0.5 Myr,
    and the box has a light crossing time of 3.3 Myr.  The shells of
    high \kph~originate from radiation that was emitted when the
    source was at its peak luminosity, illustrating the
    time-dependence of the radiative transfer equation.  Bottom:
    Radial profile \kph~with the inverse square law overplotted.}
  %\epsscale{1}
\end{figure}

\begin{figure}
  %\epsscale{0.8}
  \plotone{f31}
  \caption{\label{fig:test12_profiles} Application 4 (Time variations of the
    source luminosity).  Radial profiles of (clockwise from upper
    left) density, temperature, pressure, and ionised fraction at $t =
    50, 100, 150, 200, 250$ Myr.  In this problem, the time variations
    in the source have little effect on the overall \hii region
    expansion.  Inside the ionisation front at $t = 50$ Myr, there are
    small density perturbations that are created by the variable
    source that are later smoothed out over a sound crossing time.}
  %\epsscale{1}
\end{figure}

@@

\begin{figure*}
  \epsscale{2}
  \plotone{f32}
  \epsscale{1}
  \caption{\label{fig:test14_1} Application 5. (\hii region with MHD).
    Left to right: slices of density at $t = 0.18, 0.53, 1.58$ Myr in
    the x-y plane.  The streamlines show the magnetic field.}
\end{figure*}

\begin{figure*}
  \epsscale{2}
  \plotone{f33}
  \epsscale{1}
  \caption{\label{fig:test14_2} Application 5. (\hii region with
    MHD). Slices of density (top) and the x-component of the magnetic
    field (bottom) in the y-z plane at $t = 0.18, 0.53, 1.58$ Myr
    (left to right).}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.75}
  \plotone{f34}
  \caption{\label{fig:dx_dep1} Growth of the ionisation front radius,
    compared to the analytical radius, in Test 1 with varying spatial
    resolutions.  At resolutions of $16^3$ and $32^3$, the ionisation
    front is underestimated for the first $\sim25$ Myr but converges
    within 0.5\% of the higher resolution runs.}
  %\epsscale{1}
\end{figure}

@@

\begin{figure}
  %\epsscale{1}
  %\epsscale{2}
  \plotone{f35}
  \epsscale{1}
  \caption{\label{fig:ang_dep1} Variations in the photo-ionisation
    rates for different ray-to-cell samplings $\Phi_c$.  The colormap
    only spans a factor of 3 to enhance the contrast.  In comparison,
    the photo-ionisation rate actually spans 4 orders of magnitude in
    this test.}
  %\epsscale{1}
\end{figure}

\begin{figure}
  %\epsscale{0.75}
  \plotone{f36}
  \caption{\label{fig:ang_dep2} Standard deviations of the difference
    between the computed photo-ionisation rates and an inverse square
    law as a function of ray-to-cell samplings $\Phi_c$ for different
    spatial resolutions.  There is no dependence on the spatial
    resolution, and the accuracy increases as $\sigma \propto
    \Phi_c^{-0.6}$.}
  %\epsscale{1}
\end{figure}

@@

\begin{figure}
  %\epsscale{1.05}
  \plottwo{f37a}{f37b}
  \caption{\label{fig:nu_dep} Radial profiles of (clockwise from the
    upper left) density, temperature, radial velocity, and ionised
    fraction for Test 5 with $n_\nu = 1$, 2, 4, 8, and 16 frequency
    bins sampling the $T=10^5$ K blackbody spectrum.  The data are
    shown at $t = 200$ Myr (top) and $t = 500$ Myr (bottom).  The
    double-peaked structure in the shock only appears with a
    multi-frequency spectrum.  The solution converges at $n_\nu \ge
    4$.}
  %\epsscale{1}
\end{figure}

@@

\begin{figure}
  %\epsscale{0.75}
  \plotone{f38}
  \caption{\label{fig:dt_dep1} Growth of the ionisation front radius,
    compared to the analytical radius, in Test 1 with varying
    radiative transfer timesteps.  The $dn_{\rm H}/dt$ and $dI/dt$
    based timesteps provide the best accuracy, combined with
    computational efficiency because they take short timesteps when
    the \hii is expanding rapidly but take long timesteps when
    the photon gradients are small when $r_{\rm IF}$ is large.  At the
    final time, all but the $t = 5$ Myr constant timestep produce
    identical ionisation front radii.}
  %\epsscale{1}
\end{figure}
\begin{figure}
  %\epsscale{0.75}
  \plotone{f39}
  \caption{\label{fig:dt_dep2} Variable time-stepping for the methods
    that limit change in neutral fraction (solid) and specific
    intensity (dotted).  The horizontal lines show the constant
    timesteps that were used in the tests.  The crossing time of a
    mean free path by the ionisation front is plotted for reference.}
  %\epsscale{1}
\end{figure}

@@

\begin{figure}
  %\epsscale{0.7}
  \plotone{f40}
  \caption{\label{fig:fc_slices} Slices of the photo-ionisation rate
    in the x-y plane (top row) and x-z plane (bottom row) with (left
    column) and without (right column) the geometric correction.  The
    slices are through the origin.  In the x-y plane, it reduces the
    shell artifacts.  In the x-z plane, it reduces the severity of a
    non-spherical artifact delineated at a 45 degree angle, where the
    HEALPix scheme switches from polar to equatorial type pixels.}
  %\epsscale{1}
\end{figure}

>>>>

\begin{figure}
  \plotone{f6.eps}
  \caption{\label{fig:test1_pdf} Test 1 (\hii region expansion with a
    monochromatic spectrum of 13.6 eV). Probability distribution
    function for neutral fraction at 10 Myr (solid), 100 Myr (dashed),
    and 500 Myr (dotted).  Recombination of hydrogen at T = $10^4$ K
    causes the maximum neutral fraction to decrease from 1 to 0.75.}
\end{figure}

\begin{figure}
  \plotone{f7.eps}
  \caption{\label{fig:test1_HI} Test 1 (\hii region expansion
    with a monochromatic spectrum of 13.6 eV). Slice of neutral
    fraction at the origin at 500 Myr.}
\end{figure}

@@

\begin{figure}
  \plotone{f10}
  \caption{\label{fig:test2_pdf} Test 2. (\hii region expansion with a
    $T=10^5$ K blackbody spectrum).  Probability distribution function
    for ionized fraction (left) and temperature (right) at 10 Myr
    (solid), 100 Myr (dashed), and 500 Myr (dotted).}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{f11}
  \epsscale{1}
  \caption{\label{fig:test2_2} Test 2. (\hii region expansion
    with a $T=10^5$ K blackbody spectrum).  Top: Slices through the
    origin of neutral fraction at 10 and 100 Myr.  Bottom: Slices of
    temperature at 10 and 100 Myr.}
\end{figure*}

\begin{figure}
  \plotone{f12}
  \caption{\label{fig:test2_3} Test 2. (\hii region expansion
    with a $T=10^5$ K blackbody spectrum).  Top: Evolution of the
    radius of the simulated ionisation front (dashed) and analytical
    (solid) Str\"{o}mgren radius.  Bottom: The ratio of the calculated
    and analytical Str\"{o}mgren radius.}
\end{figure}

@@

\begin{figure}
  \plotone{f13}
  \caption{\label{fig:test3_1} Test 3. (I-front trapping in a dense
    clump and shadowing).  Line cut from the point source through the
    middle of the dense clump at $t = 1, 3, 5, 15$ Myr.  of the
    average neutral fraction (top) and temperature (bottom) of the
    clump.}
\end{figure}

\begin{figure}
  \plotone{f14}
  \caption{\label{fig:test3_2} Test 3. (I-front trapping in a dense
    clump and shadowing).  Evolution of the average ionised fraction
    (top) and temperature (bottom) of the overdense clump.}
\end{figure}

\begin{figure}
  \plotone{f15}
  \caption{\label{fig:test3_pdf} Test 3. (I-front trapping in a dense
    clump and shadowing). Probability distribution function for
    ionized fraction (left) and temperature (right) at 1 Myr (solid),
    3 Myr (dashed), and 15 Myr (dotted) within the dense clump.}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{f16}
  \epsscale{1}
  \caption{\label{fig:test3_3} Test 3. (I-front trapping in a dense
    clump and shadowing).  Clockwise from upper left: Slices through
    the origin of neutral fraction (1 Myr), temperature (1 Myr),
    temperature (15 Myr), and neutral fraction (15 Myr).}
\end{figure*}

@@

\begin{figure}
  \plotone{f17}
  \caption{\label{fig:test4_pdf} Test 4. (Multiple cosmological
    sources). Probability distribution function for neutral fraction
    (left) and temperature (right) at 50 kyr (solid) and 200 kyr
    (dashed).}
\end{figure}

\begin{figure}
  \plotone{f18}
  \caption{\label{fig:test4_1} Test 4. (Multiple cosmological
    sources).  Evolution of the mass- (dashed) and volume-averaged
    (solid) ionised fraction.}
\end{figure}

\begin{figure*}
  %\epsscale{0.9}
  \epsscale{2}
  \plotone{f19}
  \epsscale{1}
  \caption{\label{fig:test4_2} Test 4. (Multiple cosmological
    sources).  Top: Slices through the origin of neutral fraction at
    50 and 200 kyr at the coordinate $z = z_{\rm box}/2$.  Bottom:
    Slices of temperature at 50 and 200 kyr.  No smoothing has been
    applied to the images.}
  %\epsscale{1}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.8}
  \plotone{f20}
  \caption{\label{fig:test5_1} Test 5. (\hii region in a uniform
    medium).  Top: Growth of the computed ionisation front radius at
    an ionised fraction $x_e = 0.5$ (dashed) and at a temperature $T =
    10^4$ K (dotted) compared to the analytical estimate [solid;
    equation (\ref{eqn:test5_r})].  Bottom: The ratio of the computed
    ionisation front radii to the analytical estimate.} 
  %\epsscale{1}
\end{figure}

\begin{figure}
  \plotone{f21}
  \caption{\label{fig:test5_2} Test 5. (\hii region in a uniform
    medium).  Clockwise from the upper left: Radial profiles of
    density, temperature, ionised fraction, and pressure at times $t
    =$ 10, 200, and 500 Myr.}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{f22}
  \epsscale{1}
  \caption{\label{fig:test5_3} Test 5. (\hii region in a uniform
    medium).  Clockwise from the upper left: Slices through the origin
    of ionised fraction, neutral fraction, temperature, and density at
    time $t =$ 500 Myr.}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.8}
  \plotone{f23}
  \caption{\label{fig:test6_1} Test 6. (\hii region in a 1/$r^2$
    density profile).  Top: Growth of the computed ionisation front
    radius as $T=10^4$ K and $x_e = 0.5$ as definitions for the front.
    Bottom: Velocity of the ionisation front, computed from outputs at
    0.5 Myr intervals.  The velocity is calculated from $r_{\rm IF}$,
    whose coarse time resolution causes the noise seen in $v_{\rm
      IF}$.  It is smooth within the calculation itself.}
  %\epsscale{1}
\end{figure}

\begin{figure}
  \plotone{f24}
  \caption{\label{fig:test6_2} Test 6. (\hii region in a 1/$r^2$
    density profile).  Clockwise from the upper left: Radial profiles of
    density, temperature, ionised fraction, and pressure at times $t
    =$ 3, 10, and 25 Myr.}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{f25}
  \epsscale{1}
  \caption{\label{fig:test6_3} Test 6. (\hii region in a 1/$r^2$
    density profile).  Clockwise from the upper left: Slices through the origin
    of ionised fraction, neutral fraction, temperature, and density at
    time $t =$ 25 Myr.}
\end{figure*}

@@

For the first Myr, the radiation source creates a weak R-type front
where the medium is heated and ionised but does not expand because
$v_{\rm IF} > c_s$.  When $v_{\rm IF}$ becomes subsonic, the medium
can react to the passing ionisation front and creates a shock, leaving
behind a heated rarefied medium.  This behavior is clearly seen in the
radial profiles of density, temperature, ionised fraction, and
pressure in Fig. \ref{fig:test6_2}.  The inner density decreases over
two order of magnitude after 25 Myr.  To illustrate any deviations in
spherical symmetry, we show in Fig. \ref{fig:test6_3} slices of
density, temperature, neutral fraction, and ionised fraction at the
final time.  The only artifact apparent to us is the slight broadening
of the shock near the $x=0$ and $y=0$ planes.  This causes the
ionisation front radius to be slightly smaller in those directions.
In the diagonal direction, the neutral column density through the
shock is slightly smaller, allowing the high-energy photons to
photo-ionise and photo-heat the gas to $x_e = 5 \times 10^{-3}$ and $T
= 2000$ K out to $\sim50$ pc from the shock.  The reflecting
boundaries are responsible for this artifact because this is not seen
when the problem is centred in the domain, removing any boundary
effects.

@@

\begin{figure}
  %\epsscale{0.9}
  \plotone{f26}
  \caption{\label{fig:test7_1} Test 7. (Photo-evaporation of a dense
    clump).  Line cuts from the point source through the middle of the
    dense clump at $t = 1, 10, 50$ Myr of (clockwise from the upper
    left) density, temperature, pressure, and neutral fraction.}
  %\epsscale{1}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{f27}
  \epsscale{1}
  \caption{\label{fig:test7_2} Test 7. (Photo-evaporation of a dense
    clump).  Clockwise from the upper left: Slices through the clump
    centre of neutral fraction, pressure, temperature, and density at
    time $t =$ 10 Myr.}
\end{figure*}

\begin{figure*}
  \epsscale{2}
  \plotone{f28}
  \epsscale{1}
  \caption{\label{fig:test7_3} Test 7. (Photo-evaporation of a dense
    clump).  Same as Fig. \ref{fig:test7_1} but at $t = 50$ Myr.}
\end{figure*}

\begin{figure}
  \plotone{f29}
  \caption{\label{fig:test7_pdf} Test 7. (Photo-evaporation of a dense
    clump).  Probability distribution function for temperature (left)
    and flow Mach number (right) at 10 Myr (solid) and 50 Myr
    (dashed).}
\end{figure}

\begin{figure}
  \plotone{f30}
  \caption{\label{fig:test7_4} Test 7. (Photo-evaporation of a dense
    clump).  Slices through the clump centre of the flow Mach number
    at t = 10 Myr (left) and 50 Myr (right).}
\end{figure}

@@

\begin{figure*}
  \epsscale{2}
  \plotone{f31}
  \epsscale{1}
  \caption{\label{fig:test8_1} Application 1. (Champagne flow from a dense
    clump).  Slices of density through the initial clump centre in the
    x-y plane (top) and x-z plane (bottom) at $t = 10, 40, 100, 150$
    kyr.  Notice the instabilities that grow from perturbations
    created while the \hii region is contained in the dense
    clump.}
\end{figure*}

@@

The $x$ and $y$-boundaries are periodic, and the $z$-boundaries are
reflecting.  These will cause artificial features, in particular,
because of the top reflecting boundary; nevertheless, these tests
provide a stress test on a radiation hydrodynamics solver.  We show
the evolution of the density, temperature, and ionised fraction of the
optically thick and optically thin cases in Figs.
\ref{fig:test9_thick} and \ref{fig:test9_thin}.  The initial state of
the Rayleigh-Taylor instability is shown in the left panels.

@@

\begin{figure}
  %\epsscale{0.75}
  \plotone{f32a}
  \plotone{f32b}
  \plotone{f32c}
  \caption{\label{fig:test9_thick} Application 2. (Irradiated Rayleigh-Taylor
    instability; optically thick case).  Slices at $y=0$ of density
    (top), temperature (middle), and electron fraction (bottom).  The
    source turns on at $t=0$.}
  %\epsscale{1}
\end{figure}

\begin{figure}
  %\epsscale{0.75}
  \plotone{f33a}
  \plotone{f33b}
  \plotone{f33c}
  \caption{\label{fig:test9_thin} Same as Fig. \ref{fig:test9_thick}
    but for the optically thin case.}
  %\epsscale{1}
\end{figure}

@@

\begin{figure*}
  \epsscale{2}
  \plotone{f34}
  \epsscale{1}
  \caption{\label{fig:test10} Application 3 (Photo-evaporation of a
    blastwave).  Slices of density (top) and temperature (bottom) at
    $t = 2.5, 3, 5, 7.5$ Myr in the $x-z$ plane.  As the R-type
    ionisation front propagates through the blastwave centre,
    instabilities grow from the slightly inhomogeneous hot and rarefied
    medium.  Note that the dense shell of the blastwave also creates
    dense inward fingers in the ionisation front shock.}
  %\epsscale{1}
\end{figure*}

@@

\begin{figure*}
  %\epsscale{1}
  \epsscale{2}
  \plotone{f35}
  \epsscale{1}
  \caption{\label{fig:test11} Application 4 (Collimated radiation from a
    dense clump).  Slices of density (top) and temperature (bottom) at
    $t = 0.1, 3.25, 10.75, 23.25$ Myr.  The conical \hii region
    drives shocks transversely into the overdense sphere and creates
    polar champagne flows.  The ambient medium is heated to $T \sim 3
    \times 10^4$ K as the ionisation front passes the
    constant-pressure cloud surface.  The ionisation front changes
    from D-type to R-type after it enters the ambient medium.}
  %\epsscale{1}
\end{figure*}

@@

\begin{figure}
  %\epsscale{1.15}
  \plottwo{f36a}{f36b}
  \caption{\label{fig:test12_kph} Application 4 (Time variations of the
    source luminosity).  Top: Slice of the photo-ionisation rate
    \kph~through the origin.  The source has a duty cycle of 0.5 Myr,
    and the box has a light crossing time of 3.3 Myr.  The shells of
    high \kph~originate from radiation that was emitted when the
    source was at its peak luminosity, illustrating the
    time-dependence of the radiative transfer equation.  Bottom:
    Radial profile \kph~with the inverse square law overplotted.}
  %\epsscale{1}
\end{figure}

\begin{figure}
  %\epsscale{0.8}
  \plotone{f37}
  \caption{\label{fig:test12_profiles} Application 4 (Time variations of the
    source luminosity).  Radial profiles of (clockwise from upper
    left) density, temperature, pressure, and ionised fraction at $t =
    50, 100, 150, 200, 250$ Myr.  In this problem, the time variations
    in the source have little effect on the overall \hii region
    expansion.  Inside the ionisation front at $t = 50$ Myr, there are
    small density perturbations that are created by the variable
    source that are later smoothed out over a sound crossing time.}
  %\epsscale{1}
\end{figure}

@@

\begin{figure*}
  \epsscale{2}
  \plotone{f38}
  \epsscale{1}
  \caption{\label{fig:test14_1} Application 5. (\hii region with MHD).
    Left to right: slices of density at $t = 0.18, 0.53, 1.58$ Myr in
    the x-y plane.  The streamlines show the magnetic field.}
\end{figure*}

\begin{figure*}
  \epsscale{2}
  \plotone{f39}
  \epsscale{1}
  \caption{\label{fig:test14_2} Application 5. (\hii region with
    MHD). Slices of density (top) and the x-component of the magnetic
    field (bottom) in the y-z plane at $t = 0.18, 0.53, 1.58$ Myr
    (left to right).}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.75}
  \plotone{f40}
  \caption{\label{fig:dx_dep1} Growth of the ionisation front radius,
    compared to the analytical radius, in Test 1 with varying spatial
    resolutions.  At resolutions of $16^3$ and $32^3$, the ionisation
    front is underestimated for the first $\sim25$ Myr but converges
    within 0.5\% of the higher resolution runs.}
  %\epsscale{1}
\end{figure}

@@

\begin{figure}
  %\epsscale{1}
  %\epsscale{2}
  \plotone{f41}
  \epsscale{1}
  \caption{\label{fig:ang_dep1} Variations in the photo-ionisation
    rates for different ray-to-cell samplings $\Phi_c$.  The colormap
    only spans a factor of 3 to enhance the contrast.  In comparison,
    the photo-ionisation rate actually spans 4 orders of magnitude in
    this test.}
  %\epsscale{1}
\end{figure}

\begin{figure}
  %\epsscale{0.75}
  \plotone{f42}
  \caption{\label{fig:ang_dep2} Standard deviations of the difference
    between the computed photo-ionisation rates and an inverse square
    law as a function of ray-to-cell samplings $\Phi_c$ for different
    spatial resolutions.  There is no dependence on the spatial
    resolution, and the accuracy increases as $\sigma \propto
    \Phi_c^{-0.6}$.}
  %\epsscale{1}
\end{figure}

@@

\begin{figure}
  %\epsscale{1.05}
  \plottwo{f43a}{f43b}
  \caption{\label{fig:nu_dep} Radial profiles of (clockwise from the
    upper left) density, temperature, radial velocity, and ionised
    fraction for Test 5 with $n_\nu = 1$, 2, 4, 8, and 16 frequency
    bins sampling the $T=10^5$ K blackbody spectrum.  The data are
    shown at $t = 200$ Myr (top) and $t = 500$ Myr (bottom).  The
    double-peaked structure in the shock only appears with a
    multi-frequency spectrum.  The solution converges at $n_\nu \ge
    4$.}
  %\epsscale{1}
\end{figure}

@@

\begin{figure}
  %\epsscale{0.75}
  \plotone{f44}
  \caption{\label{fig:dt_dep1} Growth of the ionisation front radius,
    compared to the analytical radius, in Test 1 with varying
    radiative transfer timesteps.  The $dn_{\rm H}/dt$ and $dI/dt$
    based timesteps provide the best accuracy, combined with
    computational efficiency because they take short timesteps when
    the \hii is expanding rapidly but take long timesteps when
    the photon gradients are small when $r_{\rm IF}$ is large.  At the
    final time, all but the $t = 5$ Myr constant timestep produce
    identical ionisation front radii.}
  %\epsscale{1}
\end{figure}
\begin{figure}
  %\epsscale{0.75}
  \plotone{f45}
  \caption{\label{fig:dt_dep2} Variable time-stepping for the methods
    that limit change in neutral fraction (solid) and specific
    intensity (dotted).  The horizontal lines show the constant
    timesteps that were used in the tests.  The crossing time of a
    mean free path by the ionisation front is plotted for reference.}
  %\epsscale{1}
\end{figure}

@@

\begin{figure}
  %\epsscale{0.7}
  \plotone{f46}
  \caption{\label{fig:fc_slices} Slices of the photo-ionisation rate
    in the x-y plane (top row) and x-z plane (bottom row) with (left
    column) and without (right column) the geometric correction.  The
    slices are through the origin.  In the x-y plane, it reduces the
    shell artifacts.  In the x-z plane, it reduces the severity of a
    non-spherical artifact delineated at a 45 degree angle, where the
    HEALPix scheme switches from polar to equatorial type pixels.}
  %\epsscale{1}
\end{figure}
